Retrofitting Effect Handlers onto OCaml

### 2021-03-24 retro-concurrency.tex
use perma citations for Lwt and Async

One of the primary motivations to extend OCaml with effect handlers is to
natively support asynchronous I/O in order to express highly scalable
concurrent applications such as web servers in \emph{direct style} (as opposed
to using \emph{callbacks}). Many programming languages, including OCaml,
require non-local changes to source code in order to support asynchronous I/O,
often leading to a dichotomy between synchronous and asynchronous
code~\cite{colour}. For asynchronous I/O, OCaml developers typically use
libraries such as Lwt~\cite{lwt} and Async~\cite{async}, where asynchronous
functions are represented as monadic computations. In these libraries, while
asynchronous functions can call synchronous functions directly, the converse is
not true. In particular, any function that calls an asynchronous function will
also have to be marked as asynchronous. As a result, large parts of the
applications using these libraries end up being in monadic form. Languages such
as GHC Haskell and Go provide lightweight threads, which avoids the dichotomy
between synchronous and asynchronous code. However, these languages bake-in the
lightweight thread implementation into the runtime system. With effect
handlers, asynchronous I/O can be implemented directly in OCaml as a library
without imposing a monadic form on the users.

>>>>

One of the primary motivations to extend OCaml with effect handlers is to
natively support asynchronous I/O in order to express highly scalable
concurrent applications such as web servers in \emph{direct style} (as opposed
to using \emph{callbacks}). Many programming languages, including OCaml,
require non-local changes to source code in order to support asynchronous I/O,
often leading to a dichotomy between synchronous and asynchronous
code~\cite{colour}. For asynchronous I/O, OCaml developers typically use
libraries such as Lwt~\cite{lwt} and Async~\cite[\S18]{rwo}, where asynchronous
functions are represented as monadic computations. In these libraries, while
asynchronous functions can call synchronous functions directly, the converse is
not true. In particular, any function that calls an asynchronous function will
also have to be marked as asynchronous. As a result, large parts of the
applications using these libraries end up being in monadic form. Languages such
as GHC Haskell and Go provide lightweight threads, which avoids the dichotomy
between synchronous and asynchronous code. However, these languages bake-in the
lightweight thread implementation into the runtime system. With effect
handlers, asynchronous I/O can be implemented directly in OCaml as a library
without imposing a monadic form on the users.
